Explore gut datasets

### 2022-01-14 notebook/_posts/2021-11-09-explore-gut-data/explore-gut-data.Rmd
Add calculations to HMP2 analysis

@@

>>>>

In the HMP2 paper, they classified some samples as dysbiotic.
We don't have that info here.

@@

## Enterotypes?

Work of others suggests we might be able to identify just a handul of species/genera/groups that tend to be dominant.

```{r}

```

what taxa tend to reach dominance/very high abundance?

```{r}
ps.gen <- ps %>% tax_glom('genus')
```

```{r}
ps.gen %>%
  transform_sample_counts(close_elts) %>%
  as_tibble %>%
  filter(.abundance > 0.3) %>%
  dplyr::count(phylum, family, genus) %>%
  arrange(-n) %>%
  print(n=20)
```

```{r}
ps %>%
  tax_glom('family') %>%
  transform_sample_counts(close_elts) %>%
  as_tibble %>%
  filter(.abundance > 0.3) %>%
  dplyr::count(phylum, family) %>%
  arrange(-n) %>%
  print(n=20)
```

This suggests that when we are considering phylogenetic conservation for the purposes of bias effect, then conservation within firmicutes may not be as important as that within specific sub-groups like Lachnospiraceae and Ruminococcaceae.

Note, the original HMP protocol was found to be biased in favor of Bacteroidetes.
